Extracting Angular Observables without a Likelihood\\and Applications to Rare Decays

### 2015-03-11 draft.tex
added  scatter plots
Update: More replies to Fred, revtex4-1, PRD style (two columns)

% vim: set sts=4 et :
\documentclass[aps,prd,reprint,nofootinbib,preprintnumbers]{revtex4}

@@

Application of \refeq{mc-id} then yields
\begin{equation}
    \langle S_i\rangle \to \widehat{\langle S_i\rangle} = \frac{1}{N} \sum_{n=1}^{N} \dual{f}_i(x^{(n)})\,.
\end{equation}
It is often of interest to obtain observables integrated over certain
bins of $\nuvec$. We define
\begin{align}
    \langle S_i\rangle_{\vec{a},\vec{b}}
    & \equiv \int_{\vec{a}}^{\vec{b}} \rmdx{\nuvec} S_i(\nuvec)\\
    & = \int_{\vec{a}}^{\vec{b}} \rmdx{\nuvec} \left[\int_{\Omega} \rmdx{\vec\theta} P(\nuvec,\thvec) \dual{f}_i(\thvec)  \right]\\
    & = \int \rmdx{\nuvec} \left[\int_{\Omega} \rmdx{\vec\theta}  P(\nuvec,\thvec) \dual{f}_i(\thvec)
        \mathbf{1}(\vec{a} \le \nuvec \le \vec{b})\,
        \right],
\end{align}
where the argument of the indicator function $ \mathbf{1}(\vec{a} \le
\nuvec \le \vec{b})$ is to be interpreted componentwise.
Application of \refeq{mc-id} immediately yields
\begin{equation}
    \label{eq:bin-importance}
    \widehat{\langle S_i\rangle_{\vec{a},\vec{b}}}
    = \frac{1}{N} \sum_{n=1}^{N} \dual{f}_i(x^{(n)})         \mathbf{1}(\vec{a} \le \nuvec \le \vec{b})\,.
\end{equation}
For notational simplicity, let us forget about the $\nuvec$
integration and consider only $S_i$. In the limit $N \to \infty$, the
central limit theorem (CLT) implies that the random vector
\begin{equation}
  \label{eq:angular-obs-vec}
  \vecest{S} \equiv (\est{S_0}, \dots, \est{S_i}\,,
  \dots)
\end{equation}
follows a multivariate Gaussian distribution $\mathcal{N}(\vec{S},
\Sigma)$ centered on the true value $\vec{S}$ with the
covariance $\Sigma_{ij}$ estimated as
\begin{equation}
    \Sigma_{ij} \equiv \cov[S_i,S_j] \to \est{\Sigma_{ij}} \equiv \est{\cov}[{S}_i, {S}_j]
        = \frac{1}{N - 1} \sum_{n=1}^{N} \Big[\dual{f}_i\big(x^{(n)}\big) - \est{S_i}\Big]\,\Big[\dual{f}_j\big(x^{(n)}\big) - \widehat{S_j}\Big]\,.
\end{equation}
In our physics applications, the parameter space is compact and each
$\dual{f}_i$ is bounded. Hence the requisites for the most basic
version of the CLT to hold --- finite mean and covariance of
$\dual{f}_i$ --- are automatically satisfied.
In our numerical analysis the sample covariance rapidly converges towards the true covariance
matrix; see also \refsec{numerics}.\\

@@

\begin{figure}
    \includegraphics[width=.6\textwidth]{fig-topology.pdf}
    \caption{Decay topology for decays $B\to P_1 P_2 \ell_1 \bar\ell_2$. \label{fig:topology}}
\end{figure}

@@

For the decay at hand, we consider the partial-wave expansion for the angular observables (see \cite{Lee:1992ih} and \refapp{partial-waves})
\begin{equation}
    S_k(\vec{\nu}) \equiv \frac{I_k(\nuvec)}{4\pi \Gamma)}\,,\quad\text{with}\quad \Gamma = I_1(\nuvec) - \frac{I_2(\nuvec)}{3}\,,
\end{equation}
\fred{It not clever to use $k$ as an index and momentum variable}
\danny{I disagree. We use the same notation as in the literature, and people
should be able to distinguish these based on context.} \marcin{I am fine with k}
which reads
\begin{equation}
    \label{eq:expansion}
    S_{k}(\vec{\nu},\cos\theta_2) \equiv \sum_{j} \frac{1}{n_{l,|m|}} S_{k,l}(\vec{\nu}) p_{l}^{(|m|)}(\cos\theta_2)\sin\theta_2\,.
\end{equation}
The normalization factor $n_{l,|m|}$ is defined in \refeq{legendre-scalar-product}. Within our
example we have (cf. also \refapp{btokstarll} and \cite{Lee:1992ih})
\begin{equation}
    |m| = \begin{cases}
        0\,, & k = 1,2,6\,,\\
        1\,, & k = 4,5,7,8\,,\\
        2\,, & k = 3,9
    \end{cases}\,.
\end{equation}

@@

\fred{Ascertaining} a detector's performance to detect signal events
with accurate determination of the event's angles is generally a
difficult task. In the ideal case, one would have an explicit
probabilistic model of the detector acceptance and could thus
write down the full forward model from which the measured events
arise. In practice, that is not feasible, and one is
forced to simplify the model.
The standard approach is to generate the \emph{true} particle events
$\lbrace x_\text{t}^{(n)}\rbrace =
\lbrace(\vec{\nu}^{(n)}_\text{t},\vec\theta^{(n)}_\text{t})\rbrace$,
$n=1,\dots,N_\text{t}$, from a PDF assumed to describe the bare
physical process, and to propagate those particles through a detailed
simulation of the detector.  The observable traces that the particles
leave in the detector are fed into reconstruction algorithms resulting
in the \emph{detector} events $\lbrace x^{(n)}_\text{d}\rbrace =
\lbrace(\vec{\nu}^{(n)}_\text{d}, \vec\theta^{(n)}_\text{d})\rbrace$,
$n=1,\dots,N_\text{d}$ with $N_\text{d} \le N_\text{t}$.  In general,
the distribution of the detected events is
\begin{equation}
    P_\text{d}(x_\text{d}) = \frac{1}{R} \int P_\text{t}(x_\text{t}) E(x_\text{d} | x_\text{t}) \rmdx{x_\text{t}}\,.
\end{equation}
\fred{In this section, all integration symbols are at the end, in ch. 2 they were right after the $\int$ sign.}
Here $P_\text{t}$ is the probability distribution of the true events,
the normalization constant $R$ is given by
\begin{equation}
  \label{eq:def-R}
    R \equiv \iint P_\text{t}(x_\text{t}) E(x_\text{d} | x_\text{t}) \rmdx{x_\text{t}} \rmdx{x_\text{d}}\,.
\end{equation}
The kernel $E(x_\text{d}| x_\text{t})$ is usually decomposed as
\begin{equation}
  \Ekernel = \eps(\xt) P(\xd|\xt),
\end{equation}
where the PDF $P(\xd|\xt)$ models the resolution effects and the
unnormalized density $\eps(\xt)$ is the \emph{detector acceptance}
function. Perfect resolution corresponds to
\begin{equation}
  P(\xd|\xt) = \delta(\xd - \xt) \,.
\end{equation}

In what follows, we propose a systematic method to unfold all effects
of $\Ekernel$ through MC simulations and the method of moments,
using that $\Ekernel$ can be expanded --- at
least formally --- in Legendre polymials.  For illustration, we
proceed with the explicit example of a uniangular\footnote{%
  The generalization of this section to multiangular PDFs is
  straighforward. It can be achieved by promoting $x$ to a vector,
  promoting the Legendre polynomials to products of independent
  polynomials or spherical harmonics, and promoting the indices
  $i,j,k,m$ to multi-indices.  } PDF with $x = \cos\theta$. Let us
define the PDF in terms of the Legendre polymials (i.e., $f_k(x)
\equiv P_k(x)$) and angular observables $\vec{S}$ as
\begin{equation}
    P\subt(\xt) \equiv P\subt(\xt | \vec{S}) = \sum_k S_k f_k(\xt),
  \end{equation}
where $k = 0, 1, \dots$ denotes an angular-momentum-like index associated with the observables.
Normalization of $P\subt$ is equivalent to choosing $S_0 = 1/2$. Requiring $P\subt(\xt) \ge 0 \,\forall \xt$
implies $|S_k| \leq 1/2$ for $k > 0$. More stringent relations between the
$S_k$ might hold, but are of no concern here. For later use, we define
\begin{equation}
  \label{eq:def-Sk}
  S^{(m)}_k \equiv
  \begin{cases}
    1/2 \delta_{k,0}, & m=0\\
    1/2 (\delta_{k,0} + \delta_{k,m}), & m>0 \,,
  \end{cases}
\end{equation}
and note that
\begin{equation}
    P^{(m)}(\xt) \equiv P(\xt | \vec{S}^{(m)})
\end{equation}
is a valid PDF. The dual basis of angular functions follows then from the normalization
and orthogonality of the Legendre polynomials, and one therefore has
\begin{equation}
    \tilde{f}_k(x) = \frac{2 k + 1}{2} f_k(x),\qquad \int_{-1}^{+1} \tilde{f}_k(x) f_l(x) \rmdx{x} = \delta_{k,l}\,.
\end{equation}
We now define the \emph{simulated raw} moments $\vec{Q}^{(m)}$ as
\begin{equation}
    \label{eq:analytic-raw-moments}
    Q_i^{(m)} \equiv \iint \tilde{f}_i(\xd) P^{(m)}(\xt) \Ekernel \rmdx{\xt} \rmdx{\xd},
\end{equation}
which are instrumental to our recipe. Monte Carlo estimators of these moments can be constructed from specifically crafted detector events $x_\text{d}^{(n,m)}$, $n = 1, \dots, N_\text{d}^{(m)}$, where
\begin{equation}
    x_\text{d}^{(n,m)} \sim P_E^{(m)}(\xd) \equiv \frac{1}{R^{(m)}} \int P^{(m)}(\xt) \Ekernel \rmdx{\xt}\,.
\end{equation}
In words, for each $m$ it is required to generate events from
a toy physical distribution $P^{(m)}$, for which $S_0 = S_m=1/2$ and all other
observables are set to zero. Next, propagate these events through
a detector simulation. The normalization $R^{(m)}$ is chosen such
that $\int P_E^{(m)}(\xd) \rmdx{\xd} = 1$.  We emphasize that
$R^{(m)}$ can be estimated as $\est{R}^{(m)} = N_\text{d}^{(m)} /
N_\text{t}$, where $N_\text{t}$ corresponds to the number of
simulated true events.  The estimators then read
\begin{equation}
    \est{Q}_i^{(m)} \equiv \est{R}^{(m)} \frac{1}{N_\text{d}} \sum_n^{N_\text{d}} \tilde{f}_i(x_\text{d}^{(n,m)})
\end{equation}

Linearity of the integral over $x$ and convergence of the expansion of
$P_E^{(m)}$ in terms of Legendre polynomials ensures that
\begin{equation}
    \vec{Q}^{(m)} = M \vec{S}^{(m)}\,.
\end{equation}
We call the matrix $M$ the \emph{unfolding matrix} \fred{actually that's the folding matrix}, which is
specific to the decay at hand.
Given our definition of $S_k^{(m)}$ in
\refeq{def-Sk} it is easy to see that
\begin{equation}
    M_{ij} = \begin{cases}
        2 Q_i^{(0)}                          & j = 0\,,\\
        2\left(Q_i^{(j)} - Q_i^{(0)}\right)  & j \neq 0\,,\\
    \end{cases}
\end{equation}
and its MC estimator $\est{M}$  can be obtained through the replacements $Q_i^{(m)} \to \est{Q}_i^{(m)}$.\\

@@

%\del{The precise value of $R$ (see \refeq{def-R}) depends on the a priori
%unknown angular observables $\vec S$.  However, within ratios and
%other suitable combinations of the estimators $\vecest{S}$, the quantity
%$R$ cancels.
%Conversely, $R$ can be
%extracted from the self-consistency condition\footnote{This condition
%  is process dependent; see \refeq{recipe:lambdabtolambdall} and
%\refeq{recipe:btokstarll}.}  $\hat{S}_0 = \frac{1}{2}$. Since $R$ is
%relevant in the determination of the branching ratio, it is therefore
%useful to carry out that determination and the extraction of the
%angular observables in a combined analysis.}\\

We note that $\vecest{S}$ as determined from \refeq{estS} does not fully correspond to $\vec{S}$
for arbitrary $\eps$. Taking again the decay $B\to K\ell^+\ell^-$
as an example, \add{there are exactly three physical angular observables:} $\dim S = 3$.
Assuming an expansion of the detector acceptance $\eps$
in terms of Legendre polynomial up to a given order $L$, we have to calculate the raw moments up
to $\dim \est{Q} = \dim \est{S} = \dim S + L$. The MC estimators of the corrected angular observables
then take the following structure:
\begin{equation}
    \vecest{S} = ( \underbrace{\est{S}_0, \dots, \est{S}_{\dim S}}_\text{physical}, \underbrace{\est{S}_{\dim S + 1}, \dots, \est{S}_{\dim Q}}_\text{``superfluous''})\,.
\end{equation}
\fred{Danny: Is $\est{S}_0$ physical? I guess not. So it should be
\begin{equation}
    \vecest{S} = ( \est{S}_0, \underbrace{\est{S}_1, \dots, \est{S}_{\dim S}}_\text{physical}, \underbrace{\est{S}_{\dim S + 1}, \dots, \est{S}_{\dim Q}}_\text{``superfluous''})\,.
\end{equation}
but then $\dim S$ is not right.
}
The method is consistent as long as the MC estimators for the ``superfluous'' observables with indices $\dim S + 1$ to $\dim Q$
are compatible with zero. \footnote{This holds only for the physical model of uniangular decay distributions. If the physical
    model involves a partial-wave expansion with cutoff, these superfluous observables
correspond to higher partial waves that have been suppressed in the physical model; see \refapp{btokstarll} for such a case.}
Since $L$ depends on the setup of the particle detector under consideration, it remains for
the experimental collaborations to determine the maximum order of such an expansion.
\add{We suggest to carry out a dedicated study to determine a suitable choice for $L$.
Since such a work requires in-depth understanding of the detector simulation, it remains
beyond the scope of our possibilities.}\\

@@

We also studied the correlation of the estimators from Method of Moments to a maximum likelihood fit. We performed a toy study by simulating $10^3$ toys of 200 events. We then use for each toy method described in~\refsec{systematics:partial-waves} and a standard maximum likelihood method. In \reffig{correlation} the two estimators distributions are showed. In conclusion the two estimators are highly correlated but different between then can be of the order of statistic uncertainty of the maximum likelihood estimator.

@@

The PDF for the decay $\bar{B}\to\bar{K}\ell^+\ell^-$ has been calculated for the most
complete basis of dimension-six $b\to s \ell^+\ell^-$ operators. It reads \cite{Bobeth:2007dw,Bobeth:2012vn}
\begin{equation}
    P(q^2, \cos\theta_1)
        = \frac{1}{\rmdx{\Gamma}/\rmdx{q^2}} \frac{\rmdx{^2\Gamma}}{\rmdx{q^2} \rmdx{\cos\theta_1}}
        = \frac{a(q^2)}{\rmdx{\Gamma}/\rmdx{q^2}} + \frac{b(q^2)}{\rmdx{\Gamma}/\rmdx{q^2}} \cos\theta_1 + \frac{c(q^2)}{\rmdx{\Gamma}/\rmdx{q^2}} \cos^2\theta_1
        = \sum_i S_i p_i(\cos\theta_1)\,,
\end{equation}
with the conventional observables $a(q^2)$ through $c(q^2)$, and $\rmdx{\Gamma}/\rmdx{q^2} = 2a + 2/3 c$. We conveniently use the Legendre polynomials
$p_i(x)$, $i=0,1,2$,
\begin{equation}
\begin{aligned}
    p_0(x) & = 1\,, &
    p_1(x) & = x\,, &
    p_2(x) & = \frac{1}{2} (3x^2 - 1)\,,
\end{aligned}
\end{equation}
as our basis of angular functions. Our basis of angular observables then
translates to the conventional basis as
\begin{equation}
\begin{aligned}
    S_0 & = \frac{1}{2}\,, &
    S_1 & = \frac{b}{\Gamma}\,, &
    S_2 & = \frac{2c}{3\Gamma}\,.
\end{aligned}
\end{equation}
In this case, the dual basis is simply given by $\tilde{p}_i(x) = (2 i + 1)/2 p_i(x)$
such that
\begin{equation}
    \int_0^\pi \rmdx{\theta_1} \dual{p}_i(\cos\theta_1) P(q^2, \cos\theta_1)\sin\theta_1 = S_i(q^2)\,.
\end{equation}

@@

    P(q^2, \cos\theta_1, \cos\theta_2, \theta_3) = \frac{1}{\rmdx{\Gamma}/\rmdx{q^2}} \frac{\rmdx{^4\Gamma}}{\rmdx{q^2} \rmdx{\cos\theta_1} \rmdx{\cos\theta_2} \rmdx{\theta_3}}
    = \sum_i S_i(q^2) f_i(\cos\theta_1, \cos\theta_2, \theta_3)\,,

@@

We use the same bases of angular functions as given in \refeq{lambdab:bases} for the decay $\Lambda_b\to \Lambda\ell^+\ell^-$.

@@

    8\pi S_{2, 2, -2} & = \frac{8 J_9}{\rmdx{\Gamma}/\rmdx{q^2}}                                           \\

@@

    8\pi S_{2, 2, +2} & = \frac{8 J_3}{\rmdx{\Gamma}/\rmdx{q^2}}                                           \\

@@

As before, for the purpose of unfolding acceptance effects as laid down in \refsec{systematics:acceptance}, it is instrumental
to know that $f_{0,0,0} \equiv 1$, and that $\max_{\cos\theta_1,\cos\theta_2,\phi} |f_{l_1, l_2, m}| \leq 1$.
The recipe's generating PDFs are therefore
\begin{equation}
    \label{eq:recipe:btokstarll}
    P(x|\lbrace S_i^{(j)}\rbrace)\,,\quad\text{with}\quad S_i^{(j)} = \frac{1}{8\pi}
        \begin{cases}
            \delta_{i,(0,0,0)}                  & j = (0, 0, 0)\\
            \delta_{i,(0,0,0)}  + \delta_{i,j}  & j \neq (0, 0, 0)
        \end{cases}
\end{equation}
where also $j$ is now a multi-index representing $j \sim (l_1, l_2, m)$.

>>>>

% vim: set sts=4 et :
\documentclass[aps,nofootinbib,preprintnumbers,prd,twocolumn]{revtex4-1}

@@

Application of \refeq{mc-id} then yields
\begin{equation}
    \langle S_i\rangle \to \widehat{\langle S_i\rangle} = \frac{1}{N} \sum_{n=1}^{N} \dual{f}_i(x^{(n)})\,.
\end{equation}
It is often of interest to obtain observables integrated over certain
bins of $\nuvec$. We define
\begin{align}
    \langle S_i\rangle_{\vec{a},\vec{b}}
    & \equiv \int_{\vec{a}}^{\vec{b}} \rmdx{\nuvec} S_i(\nuvec)\\
    & = \int_{\vec{a}}^{\vec{b}} \rmdx{\nuvec} \left[\int_{\Omega} \rmdx{\vec\theta} P(\nuvec,\thvec) \dual{f}_i(\thvec)  \right]\\
    & = \int \rmdx{\nuvec} \left[\int_{\Omega} \rmdx{\vec\theta}  P(\nuvec,\thvec) \dual{f}_i(\thvec)
        \mathbf{1}(\vec{a} \le \nuvec \le \vec{b})\,
        \right],
\end{align}
where the argument of the indicator function $ \mathbf{1}(\vec{a} \le
\nuvec \le \vec{b})$ is to be interpreted componentwise.
Application of \refeq{mc-id} immediately yields
\begin{equation}
    \label{eq:bin-importance}
    \widehat{\langle S_i\rangle_{\vec{a},\vec{b}}}
    = \frac{1}{N} \sum_{n=1}^{N} \dual{f}_i(x^{(n)})         \mathbf{1}(\vec{a} \le \nuvec \le \vec{b})\,.
\end{equation}
For notational simplicity, let us forget about the $\nuvec$
integration and consider only $S_i$. In the limit $N \to \infty$, the
central limit theorem (CLT) implies that the random vector
\begin{equation}
  \label{eq:angular-obs-vec}
  \vecest{S} \equiv (\est{S_0}, \dots, \est{S_i}\,,
  \dots)
\end{equation}
follows a multivariate Gaussian distribution $\mathcal{N}(\vec{S},
\Sigma)$ centered on the true value $\vec{S}$ with the
covariance $\Sigma_{ij}$ estimated as
\begin{equation}
\begin{aligned}
    \Sigma_{ij}
        & \equiv \cov[S_i,S_j]\\
        & \to \est{\Sigma_{ij}} \equiv \est{\cov}[{S}_i, {S}_j]\\
        & = \frac{1}{N - 1} \sum_{n=1}^{N} \Big[\dual{f}_i\big(x^{(n)}\big) - \est{S_i}\Big]\,\Big[\dual{f}_j\big(x^{(n)}\big) - \widehat{S_j}\Big]\,.
\end{aligned}
\end{equation}
In our physics applications, the parameter space is compact and each
$\dual{f}_i$ is bounded. Hence the requisites for the most basic
version of the CLT to hold --- finite mean and covariance of
$\dual{f}_i$ --- are automatically satisfied.
In our numerical analysis the sample covariance rapidly converges towards the true covariance
matrix; see also \refsec{numerics}.\\

@@

\begin{figure}
    \includegraphics[width=.45\textwidth]{fig-topology.pdf}
    \caption{Decay topology for decays $B\to P_1 P_2 \ell_1 \bar\ell_2$. \label{fig:topology}}
\end{figure}

@@

For the decay at hand, we consider the partial-wave expansion for the angular observables (see \cite{Lee:1992ih} and \refapp{partial-waves})
\begin{equation}
    S_k(\vec{\nu}) \equiv \frac{I_k(\nuvec)}{4\pi \Gamma)}\,,\quad\text{with}\quad \Gamma = I_1(\nuvec) - \frac{I_2(\nuvec)}{3}\,,
\end{equation}
\fred{It not clever to use $k$ as an index and momentum variable}
\danny{I disagree. We use the same notation as in the literature, and people
should be able to distinguish these based on context.} \marcin{I am fine with k}
which reads
\begin{multline}
    \label{eq:expansion}
    S_{k}(\vec{\nu},\cos\theta_2) \\
        \equiv \sum_{j} \frac{1}{n_{l,|m|}} S_{k,l}(\vec{\nu}) p_{l}^{(|m|)}(\cos\theta_2)\sin\theta_2\,.
\end{multline}
The normalization factor $n_{l,|m|}$ is defined in \refeq{legendre-scalar-product}. Within our
example we have (cf. also \refapp{btokstarll} and \cite{Lee:1992ih})
\begin{equation}
    |m| = \begin{cases}
        0\,, & k = 1,2,6\,,\\
        1\,, & k = 4,5,7,8\,,\\
        2\,, & k = 3,9
    \end{cases}\,.
\end{equation}

@@

Ascertaining a detector's performance to detect signal events
with accurate determination of the event's angles is generally a
difficult task. In the ideal case, one would have an explicit
probabilistic model of the detector acceptance and could thus
write down the full forward model from which the measured events
arise. In practice, that is not feasible, and one is
forced to simplify the model.
The standard approach is to generate the \emph{true} particle events
$\lbrace x_\text{t}^{(n)}\rbrace =
\lbrace(\vec{\nu}^{(n)}_\text{t},\vec\theta^{(n)}_\text{t})\rbrace$,
$n=1,\dots,N_\text{t}$, from a PDF assumed to describe the bare
physical process, and to propagate those particles through a detailed
simulation of the detector.  The observable traces that the particles
leave in the detector are fed into reconstruction algorithms resulting
in the \emph{detector} events $\lbrace x^{(n)}_\text{d}\rbrace =
\lbrace(\vec{\nu}^{(n)}_\text{d}, \vec\theta^{(n)}_\text{d})\rbrace$,
$n=1,\dots,N_\text{d}$ with $N_\text{d} \le N_\text{t}$.  In general,
the distribution of the detected events is
\begin{equation}
    P_\text{d}(x_\text{d}) = \frac{1}{R} \int \rmdx{x_\text{t}} P_\text{t}(x_\text{t}) E(x_\text{d} | x_\text{t}) \,.
\end{equation}
\fred{In this section, all integration symbols are at the end, in ch. 2 they were right after the $\int$ sign.}
Here $P_\text{t}$ is the probability distribution of the true events,
the normalization constant $R$ is given by
\begin{equation}
  \label{eq:def-R}
    R \equiv \iint \rmdx{x_\text{t}} \rmdx{x_\text{d}} P_\text{t}(x_\text{t}) E(x_\text{d} | x_\text{t})\,.
\end{equation}
The kernel $E(x_\text{d}| x_\text{t})$ is usually decomposed as
\begin{equation}
  \Ekernel = \eps(\xt) P(\xd|\xt),
\end{equation}
where the PDF $P(\xd|\xt)$ models the resolution effects and the
unnormalized density $\eps(\xt)$ is the \emph{detector acceptance}
function. Perfect resolution corresponds to
\begin{equation}
  P(\xd|\xt) = \delta(\xd - \xt) \,.
\end{equation}

In what follows, we propose a systematic method to unfold all effects
of $\Ekernel$ through MC simulations and the method of moments,
using that $\Ekernel$ can be expanded --- at
least formally --- in Legendre polymials.  For illustration, we
proceed with the explicit example of a uniangular\footnote{%
  The generalization of this section to multiangular PDFs is
  straighforward. It can be achieved by promoting $x$ to a vector,
  promoting the Legendre polynomials to products of independent
  polynomials or spherical harmonics, and promoting the indices
  $i,j,k,m$ to multi-indices.  } PDF with $x = \cos\theta$. Let us
define the PDF in terms of the Legendre polymials (i.e., $f_k(x)
\equiv P_k(x)$) and angular observables $\vec{S}$ as
\begin{equation}
    P\subt(\xt) \equiv P\subt(\xt | \vec{S}) = \sum_k S_k f_k(\xt),
  \end{equation}
where $k = 0, 1, \dots$ denotes an angular-momentum-like index associated with the observables.
Normalization of $P\subt$ is equivalent to choosing $S_0 = 1/2$. Requiring $P\subt(\xt) \ge 0 \,\forall \xt$
implies $|S_k| \leq 1/2$ for $k > 0$. More stringent relations between the
$S_k$ might hold, but are of no concern here. For later use, we define
\begin{equation}
  \label{eq:def-Sk}
  S^{(m)}_k \equiv
  \begin{cases}
    1/2 \delta_{k,0}, & m=0\\
    1/2 (\delta_{k,0} + \delta_{k,m}), & m>0 \,,
  \end{cases}
\end{equation}
and note that
\begin{equation}
    P^{(m)}(\xt) \equiv P(\xt | \vec{S}^{(m)})
\end{equation}
is a valid PDF. The dual basis of angular functions follows then from the normalization
and orthogonality of the Legendre polynomials, and one therefore has
\begin{gather}
    \tilde{f}_k(x) = \frac{2 k + 1}{2} f_k(x)\,,\\
    \int_{-1}^{+1} \rmdx{x} \tilde{f}_k(x) f_l(x) = \delta_{k,l}\,.
\end{gather}
We now define the \emph{simulated raw} moments $\vec{Q}^{(m)}$ as
\begin{equation}
    \label{eq:analytic-raw-moments}
    Q_i^{(m)} \equiv \iint \rmdx{\xt} \rmdx{\xd} \tilde{f}_i(\xd) P^{(m)}(\xt) \Ekernel,
\end{equation}
which are instrumental to our recipe. Monte Carlo estimators of these moments can be constructed from specifically crafted detector events $x_\text{d}^{(n,m)}$, $n = 1, \dots, N_\text{d}^{(m)}$, where
\begin{equation}
    x_\text{d}^{(n,m)} \sim P_E^{(m)}(\xd) \equiv \frac{1}{R^{(m)}} \int \rmdx{\xt} P^{(m)}(\xt) \Ekernel\,.
\end{equation}
In words, for each $m$ it is required to generate events from
a toy physical distribution $P^{(m)}$, for which $S_0 = S_m=1/2$ and all other
observables are set to zero. Next, propagate these events through
a detector simulation. The normalization $R^{(m)}$ is chosen such
that $\int P_E^{(m)}(\xd) \rmdx{\xd} = 1$.  We emphasize that
$R^{(m)}$ can be estimated as $\est{R}^{(m)} = N_\text{d}^{(m)} /
N_\text{t}$, where $N_\text{t}$ corresponds to the number of
simulated true events.  The estimators then read
\begin{equation}
    \est{Q}_i^{(m)} \equiv \est{R}^{(m)} \frac{1}{N_\text{d}} \sum_n^{N_\text{d}} \tilde{f}_i(x_\text{d}^{(n,m)})
\end{equation}

Linearity of the integral over $x$ and convergence of the expansion of
$P_E^{(m)}$ in terms of Legendre polynomials ensures that
\begin{equation}
    \vec{Q}^{(m)} = M \vec{S}^{(m)}\,.
\end{equation}
We call the matrix \add{$M^{-1}$} the \emph{unfolding matrix},
\fred{actually that's the folding matrix}
\danny{switched to $M^{-1}$ instead. OK?}
which is specific to the decay at hand.
Given our definition of $S_k^{(m)}$ in
\refeq{def-Sk} it is easy to see that
\begin{equation}
    M_{ij} = \begin{cases}
        2 Q_i^{(0)}                          & j = 0\,,\\
        2\left(Q_i^{(j)} - Q_i^{(0)}\right)  & j \neq 0\,,\\
    \end{cases}
\end{equation}
and its MC estimator $\est{M}$  can be obtained through the replacements $Q_i^{(m)} \to \est{Q}_i^{(m)}$.\\

@@

We note that $\vecest{S}$ as determined from \refeq{estS} does not fully correspond to $\vec{S}$
for arbitrary $\eps$. Taking again the decay $B\to K\ell^+\ell^-$
as an example, \add{there are exactly three physical angular observables:} $\dim S = 3$.
Assuming an expansion of the detector acceptance $\eps$
in terms of Legendre polynomial up to a given order $L$, we have to calculate the raw moments up
to $\dim \est{Q} = \dim \est{S} = \dim S + L$. The MC estimators of the corrected angular observables
then take the following structure:
\begin{equation}
    \vecest{S} = ( \underbrace{\est{S}_0, \dots, \est{S}_{\dim S}}_\text{physical}, \underbrace{\est{S}_{\dim S + 1}, \dots, \est{S}_{\dim Q}}_\text{``superfluous''})\,.
\end{equation}
\fred{Danny: Is $\est{S}_0$ physical? I guess not. So it should be
\begin{equation}
    \vecest{S} = ( \est{S}_0, \underbrace{\est{S}_1, \dots, \est{S}_{\dim S}}_\text{physical}, \underbrace{\est{S}_{\dim S + 1}, \dots, \est{S}_{\dim Q}}_\text{``superfluous''})\,.
\end{equation}
but then $\dim S$ is not right.
}
\danny{Fred: We simplfy define the label ``physical'' right here, which includes
$S_0$. Although perhaps a better label would be ``nominal''?}
The method is consistent as long as the MC estimators for the ``superfluous'' observables with indices $\dim S + 1$ to $\dim Q$
are compatible with zero. \footnote{This holds only for the physical model of uniangular decay distributions. If the physical
    model involves a partial-wave expansion with cutoff, these superfluous observables
correspond to higher partial waves that have been suppressed in the physical model; see \refapp{btokstarll} for such a case.}
Since $L$ depends on the setup of the particle detector under consideration, it remains for
the experimental collaborations to determine the maximum order of such an expansion.
\add{We suggest to carry out a dedicated study to determine a suitable choice for $L$.
Since such a work requires in-depth understanding of the detector simulation, it remains
beyond the scope of our possibilities.}\\

@@

We also studied the correlation of the estimators from Method of Moments to a maximum likelihood fit. We performed a toy study by simulating $10^3$ toys of 200 events. We then use for each toy method described in~\refsec{systematics:partial-waves} and a standard maximum likelihood method. In \reffig{correlation} the two estimators distributions are showed. In conclusion the two estimators are highly correlated but different between then can be of the order of statistic uncertainty of the maximum likelihood estimator. We found that in some observables this can reach almost half of the statistical error.

@@

The PDF for the decay $\bar{B}\to\bar{K}\ell^+\ell^-$ has been calculated for the most
complete basis of dimension-six $b\to s \ell^+\ell^-$ operators. It reads \cite{Bobeth:2007dw,Bobeth:2012vn}
\begin{equation}
\begin{aligned}
    P(q^2, \cos\theta_1)
        & = \frac{1}{\rmdx{\Gamma}/\rmdx{q^2}} \frac{\rmdx{^2\Gamma}}{\rmdx{q^2} \rmdx{\cos\theta_1}}\\
        & = \frac{a(q^2)}{\rmdx{\Gamma}/\rmdx{q^2}} + \frac{b(q^2)}{\rmdx{\Gamma}/\rmdx{q^2}} \cos\theta_1\\
        & \quad + \frac{c(q^2)}{\rmdx{\Gamma}/\rmdx{q^2}} \cos^2\theta_1\\
        & \equiv \sum_i S_i p_i(\cos\theta_1)\,,
\end{aligned}
\end{equation}
with the conventional observables $a(q^2)$ through $c(q^2)$, and $\rmdx{\Gamma}/\rmdx{q^2} = 2a + 2/3 c$. We conveniently use the Legendre polynomials
$p_i(x)$, $i=0,1,2$,
\begin{equation}
\begin{aligned}
    p_0(x) & = 1\,, &
    p_1(x) & = x\,, &
    p_2(x) & = \frac{1}{2} (3x^2 - 1)\,,
\end{aligned}
\end{equation}
as our basis of angular functions. Our basis of angular observables then
translates to the conventional basis as
\begin{equation}
\begin{aligned}
    S_0 & = \frac{1}{2}\,, &
    S_1 & = \frac{b}{\Gamma}\,, &
    S_2 & = \frac{2c}{3\Gamma}\,.
\end{aligned}
\end{equation}
In this case, the dual basis is simply given by $\tilde{p}_i(x) = (2 i + 1)/2 p_i(x)$
such that
\begin{equation}
    \int_0^\pi \rmdx{\theta_1} \dual{p}_i(\cos\theta_1) P(q^2, \cos\theta_1)\sin\theta_1 = S_i(q^2)\,.
\end{equation}

@@

    P(q^2, \cos\theta_1, \cos\theta_2, \theta_3)
    & = \frac{(\rmdx{\Gamma}/\rmdx{q^2})^{-1}\rmdx{^4\Gamma}}{\rmdx{q^2} \rmdx{\cos\theta_1} \rmdx{\cos\theta_2} \rmdx{\theta_3}}\\
    & = \sum_i S_i(q^2) f_i(\cos\theta_1, \cos\theta_2, \theta_3)\,,

@@

We use the same bases of angular functions as given in \refeq{lambdab:bases1} and \refeq{lambdab:bases2}
for the decay $\Lambda_b\to \Lambda\ell^+\ell^-$.
\begin{widetext}

@@

    8\pi S_{2, 2, -2} & = \frac{8 J_9}{\rmdx{\Gamma}/\rmdx{q^2}}\,,                                        \\

@@

    8\pi S_{2, 2, +2} & = \frac{8 J_3}{\rmdx{\Gamma}/\rmdx{q^2}}\,.                                        \\

@@

\end{widetext}

As before, for the purpose of unfolding acceptance effects as laid down in \refsec{systematics:acceptance}, it is instrumental
to know that $f_{0,0,0} \equiv 1$, and that $\max_{\cos\theta_1,\cos\theta_2,\phi} |f_{l_1, l_2, m}| \leq 1$.
The recipe's generating PDFs are therefore $P(x|\lbrace S_i^{(j)}\rbrace)$ with
\begin{equation}
    \label{eq:recipe:btokstarll}
    S_i^{(j)} = \frac{1}{8\pi}
        \begin{cases}
            \delta_{i,(0,0,0)}                  & j = (0, 0, 0)\\
            \delta_{i,(0,0,0)}  + \delta_{i,j}  & j \neq (0, 0, 0)
        \end{cases}
\end{equation}
where also $j$ is now a multi-index representing $j \sim (l_1, l_2, m)$.
